\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-10-03 gwnnDraft02.tex
small edits

Gravitational waves are produced with contributions from different components of the SFOPT's evolution.  It is commonplace to parameterize the spectral energy density in gravitational waves by~\citep{PhysRevD.75.043507} 
\begin{equation}
\Omega_{\textrm{GW}} (f) \equiv \frac{1}{ \rho_{c}} \frac{d \rho_{\textrm{GW}}(f) }{d \textrm{\,log}(f)} , 
\end{equation}
where $\rho_{c} = 3H^2/(8 \pi G)$ is the critical energy density. The total gravitational wave signal is a linear combination of three leading contributions:
\begin{equation}
h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi} + h^2\Omega_{v} + h^2\Omega_{turb} .
\end{equation}
In essence, this is a linear combination of the individual spectral shapes weighted by their efficiency factors $\kappa$. 
The three leading order contributions to the GW power spectrum are as follows:
\begin{itemize}
\item \textbf{Scalar field contributions $\Omega_{\phi}$}: Caused by collisions of the bubble walls, the solutions being completely dependent on the scalar field configuration. With efficiency factor $\kappa_{\phi} = 1 - \alpha_{\infty}/\alpha$~\citep{PhysRevD.45.4514, Huber_2008}. 
\item \textbf{Sound wave contributions $\Omega_{v}$}: Sound waves within the plasma after bubble collision will produce $\beta/H$ enhanced gravitational waves. With efficiency factor $\kappa_{v} \propto \alpha_{\infty}/\alpha$~\citep{PhysRevLett.112.041301}.
\item \textbf{Magnetohydrodynamical contributions $\Omega_{B}$}: Turbulence within the plasma, left over from the sound wave propagation, will produce gravitational waves. With efficiency factor $\kappa_{
turb} \approx 0.1 \kappa_{v} $~\citep{PhysRevD.74.063521}.
\end{itemize}
The parameter $\alpha_{\infty}$ denotes the dividing line between the runaway regime  $(\alpha >\alpha_{\infty})$ and the non-runaway regime $(\alpha <\alpha_{\infty})$. Explicitly \cite{Breitbach:2018ddu, Caprini:2015zlo, Espinosa:2010hh}, 
\begin{equation}\label{eqn:critPTstrength}
\alpha_{\infty} = \frac{(T^{nuc}_h)^2}{\rho_R}\left[\sum_{bosons} n_i\frac{\Delta m^2_i}{24} + \sum_{fermions} n_i\frac{\Delta m_i^2}{48}\right],
\end{equation} 
for particles with $n_i$ degrees of freedom that obtain mass through the phase transition. The exotic sectors have massless degrees of freedom pre-phase transition and pions and baryons with negligible masses post-phase transition, leading to small $\alpha_{\infty}$ from Eq.~(\ref{eqn:critPTstrength}). \footnote {It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$.}. Heavy standard sectors that undergo SFOPT for QCD feature no baryons due to all quarks being above their respective QCD scales and, as such, produce runaway bubbles similar to the exotic sectors in standard $N$naturalness.Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Thus, all exotic sectors have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). In this case, the efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$.

@@

This is not the case if we consider more generalized reheating scenarios. Once the restriction that sectors with small Higgs masses are preferentially reheated has been lifted, we can explore a much more vast landscape of hidden sectors than are allowed in the vanilla case. Here, we construct several different scenarios that are both detectable and demonstrate a variety of gravitational wave profiles. Specifically, we explore scenarios that lead to a deviation in the peak behaviour of the total GW signal (the superposition of stochastic GW from individual SFOPT) from a standard power law signal.

@@

In this work, it is assumed that $\delta t = 14$ days, but this may underestimate the capabilities of SKA as well as the cadences of the pulsar populations. If the timing residuals are lowered the maximum frequency reach of SKA increases, and the corresponding PSI curves in Figs.~\ref{fig::Nnatural} $\&$ \ref{fig:Haa} are shifted to the right, possibly giving the PTAs sensitivity to some of the scenarios considered here.

@@

Here, we examined scenarios including $N$naturalness that involve multiple hidden sectors and calculated the GW profiles present. Our GW projections demonstrate that although vanilla $N$naturalness is not projected to be detectable in the near future, more generalized scenarios with multiple hidden sector SFOPTs are in an observable region and will begin to be probed by next generation space experiments. Both cases feature important parts of their GW signals in the void between frequencies detectable by pulsar timing arrays and space-based interferometers --- providing theoretical impetus for new experiments capable of probing this region of frequency space.

>>>>

Gravitational waves are produced with contributions from different components of the SFOPT's evolution.  It is commonplace to parameterize the spectral energy density in gravitational waves by~\citep{PhysRevD.75.043507} 
\begin{equation}
\Omega_{\textrm{GW}} (f) \equiv \frac{1}{ \rho_{c}} \frac{d \rho_{\textrm{GW}}(f) }{d \textrm{\,log}(f)} , 
\end{equation}
where $\rho_{c} = 3H^2/(8 \pi G)$ is the critical energy density. The total gravitational wave signal is a linear combination of three leading contributions:
\begin{equation}
h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi} + h^2\Omega_{v} + h^2\Omega_{turb} .
\end{equation}
In essence, this is a linear combination of the individual spectral shapes weighted by their efficiency factors $\kappa$. 
The three leading order contributions to the GW power spectrum are as follows:
\begin{itemize}
\item \textbf{Scalar field contributions $\Omega_{\phi}$}: Caused by collisions of the bubble walls, the solutions being completely dependent on the scalar field configuration. With efficiency factor $\kappa_{\phi} = 1 - \alpha_{\infty}/\alpha$~\citep{PhysRevD.45.4514, Huber_2008}. 
\item \textbf{Sound wave contributions $\Omega_{v}$}: Sound waves within the plasma after bubble collision will produce $\beta/H$ enhanced gravitational waves. With efficiency factor $\kappa_{v} \propto \alpha_{\infty}/\alpha$~\citep{PhysRevLett.112.041301}.
\item \textbf{Magnetohydrodynamical contributions $\Omega_{B}$}: Turbulence within the plasma, left over from the sound wave propagation, will produce gravitational waves. With efficiency factor $\kappa_{
turb} \approx 0.1 \kappa_{v} $~\citep{PhysRevD.74.063521}.
\end{itemize}
The parameter $\alpha_{\infty}$ denotes the dividing line between the runaway regime  $(\alpha >\alpha_{\infty})$ and the non-runaway regime $(\alpha <\alpha_{\infty})$. Explicitly \cite{Breitbach:2018ddu, Caprini:2015zlo, Espinosa:2010hh}, 
\begin{equation}\label{eqn:critPTstrength}
\alpha_{\infty} = \frac{(T^{nuc}_h)^2}{\rho_R}\left[\sum_{bosons} n_i\frac{\Delta m^2_i}{24} + \sum_{fermions} n_i\frac{\Delta m_i^2}{48}\right],
\end{equation} 
for particles with $n_i$ degrees of freedom that obtain mass through the phase transition. The exotic sectors have massless degrees of freedom pre-phase transition and pions and baryons with negligible masses post-phase transition, leading to small $\alpha_{\infty}$ from Eq.~(\ref{eqn:critPTstrength})\footnote {It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$.}. Heavy standard sectors that undergo SFOPT for QCD feature no baryons due to all quarks being above their respective QCD scales and, as such, produce runaway bubbles similar to the exotic sectors in standard $N$naturalness. Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Explicitly, all exotic sectors have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). In this case, the efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$.

@@

This is not the case if we consider more generalized reheating scenarios. Once the restriction that sectors with small Higgs masses are preferentially reheated has been lifted, we can explore a much more vast landscape of hidden sectors than are allowed in the vanilla case. Here, we construct several different scenarios that are both detectable and demonstrate a variety of gravitational wave profiles. Specifically, we explore benchmarks that lead to a deviation in the peak behaviour of the total GW signal (the superposition of stochastic GW from individual SFOPT) from a standard power law signal.

@@

In this work, it is assumed that $\delta t = 14$ days, but this may underestimate the capabilities of SKA as well as the cadences of the pulsar populations. If the timing residuals are lowered the maximum frequency reach of SKA increases, and the corresponding PLI curves in Figs.~\ref{fig::Nnatural} $\&$ \ref{fig:Haa} are shifted to the right, possibly giving the PTAs sensitivity to some of the scenarios considered here.

@@

Here, we examined scenarios, including $N$naturalness, that involve multiple hidden sectors and calculated the GW profiles present. Our GW projections demonstrate that although vanilla $N$naturalness is not projected to be detectable in the near future, more generalized scenarios with multiple hidden sector SFOPTs are in an observable region and will begin to be probed by next generation space experiments. Both cases feature important parts of their GW signals in the void between frequencies detectable by pulsar timing arrays and space-based interferometers --- providing theoretical impetus for new experiments capable of probing this region of frequency space.
